---
title: "Spirit Score Analysis"
author: "csaund"
date: "11/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggplot2)
library(pwr)
library(ordinal)
library(gridExtra)
library(RColorBrewer)
library(knitr)
library(scales)
library(googledrive)
```

I couldn't get loading from a google drive pack working, so I copied loads from here:
https://docs.google.com/spreadsheets/d/e/2PACX-1vSpxvl3g-EBCqR96hJ83V5pTN68k88cKiDF9eMgVP6zjDnn_El8emYfUQ2epBq-uu8yTtW37GzIV1pG/pubhtml#. 
If you're also an engineer, it makes your blood boil that I didn't automate this. Big shrug.

Honestly, I can only describe UKU's archival data mangement and reporting as a flaming garbage heap. A totally understandable flaming garbage heap, but a complete and utter dumpster fire nonetheless. EUCF and WFDF are no better, to be clear. Overall, actually getting this data was by far the most time-consuming and least rewarding part of this process. Before embarking I had considered trying to put together a little program that would scrape and visualize scores from these three sources and make a little app where you could filter by league, age, division, etc. and it would create visualizations to compare with, but the headache of getting the data is just not worth it right now, cause I have a real job already where I code all day. But if you want this app and you want to pay me to take a sebatical from my PhD to make it, talk to me, cause that is 100% something I would be down to do.


Consequently, the super-caveat to all of this is that this is that there's not a ton of data to work with. Our overall dataset is pretty large, but we're often aggregating indoor and outdoor, and often looking across multiple years to answer some of our questions. Worse yet, our mixed data all comes from indoor, but our Men's/Women's data is mixed across indoor and outdoor. Also, at the moment this is all about *UK, University-Level Ultimate* (for aforementioned data-collection reasons). Basically what I'm saying is, this is all very wishy-washy and about general trends, because if we actually controlled for everything we wouldn't have enough data to see interesting trends. If you provide me with spreadsheets of other regions, I will happily run some of these comparisons on them!



First we load in the data that I comglomerated because R doesn't know how to read in multiple spreadsheets from google drive. 
```{r}
spirit_path = "spirit.csv"
dl <- drive_download(as_id("1wmn1K04N4dRuk4HPezT0fHFl0q1pEeSUsjgm2hfpXPc"), path=spirit_path, overwrite=TRUE)
# read in and add total for ease
raw <- read.csv(spirit_path) %>%
  mutate(total = Rules + Fouls + Fairness + Attitude + Comms) %>%
  mutate(original_index=as.numeric(rownames(raw))) %>%
  na.omit()

# I feel like the above shouldn't work and I should have to do the below
# raw <- mutate(raw, original_index=as.numeric(rownames(raw)))


## just get all these set up for later.
women <- raw %>%
  filter(League == "Women") %>%
  na.omit()

men <- raw %>%
  filter(League == "Men") %>%
  na.omit()

mixed <- raw %>%
  filter(League == "Mixed") %>%
  na.omit()

div.i <- raw %>%
  filter(Level == "Nats Div I")

div.ii <- raw %>%
  filter(Level == "Nats Div II")

regionals <- raw %>%
  filter(Level == "Regionals")

indoor <- raw %>%
  filter(Indoor.Outdoor == "Indoor")

outdoor <- raw %>%
  filter(Indoor.Outdoor == "Outdoor")

repeats <- raw %>%
  filter(Repeat == "Repeat Game")
```

Now let's just start asking questions because data science is dope AF. 

1. Are spirit scores reciprocal? That is, given one team thought the spirit was one way, did the other team also think that?

2. Which teams have beef with each other? 

3. How do repeat games stack up against one another?
```{r}
# because of the alphabet I was very cheeky in how I discerned "repeat" and "first" games
repeat_indexes <- repeats$original_index
first_game_indexes <- repeat_indexes - 1

first_games <- raw %>%
  filter(original_index %in% first_game_indexes)

#stats
first_game_avg <- first_games$total
repeat_game_avg <- repeats$total

# first game stats:
mean(first_game_avg)
sd(first_game_avg)

# repeat game stats:
mean(repeat_game_avg)
sd(repeat_game_avg)
```
Almost no difference


Plot them now
```{r}
p <- ggplot(data=df2, aes(x=dose, y=len, fill=Repeat)) +
geom_bar(stat="identity", color="black", position=position_dodge())+
  theme_minimal()
# Use custom colors
p + scale_fill_manual(values=c('#999999','#E69F00'))
# Use brewer color palettes
p + scale_fill_brewer(palette="Blues")

## there are many more clever says to do this, I am just in a pinch for time
w_att <- mean(women$Attitude)
m_att <- mean(men$Attitude)
x_att <- mean(mixed$Attitude)

df <- data.frame("score"=c(w_att, m_att, x_att), "league"=c("women", "men", "mixed"), "category"=rep("attitude", 3))


plotable_data <- raw %>% 
  gather("category", "score", 5:9) 

p <- ggplot(plotable_data) + 
  geom_bar(aes(x=category, y=score, fill=League), stat="identity", alpha=0.7) + 
  geom_errorbar(aes(x=category, ymin=score-sd, ymax=score+sd), width=0.4, alpha=0.9)
  
p <- ggplot(data=df, aes(x=category, y=score, fill=league)) + 
  geom_bar(stat="identity", color="black", position=position_dodge())

p


summarise_attr <- function(groupby, attr) {
  d <- raw %>%
    na.omit() %>%
    group_by(League) %>%
    summarise(mean_attr=mean(attr),
              sd_attr=sd(attr),
              n_gr=n(),
              se_attr=sd(attr)/sqrt(n()))
  return (d)
}

d<-summarise_attr("League", "Rules")

t <- raw %>%
  na.omit() %>%
  group_by(League)

rules_summary <- raw %>%
  na.omit() %>%
  group_by(League) %>%
  summarise(mean_rules=mean(Rules),
            sd_rules=sd(Rules),
            n_Gr=n(),
            SE_rules=sd(Rules)/sqrt(n()))

attitude_summary <- raw %>%
  na.omit() %>%
  group_by(League) %>%
  summarise(mean_attitude=mean(Attitude),
            sd_attitude=sd(Attitude),
            n_Gr=n(),
            SE_attitude=sd(Attitude)/sqrt(n()))


p <- ggplot(rules_summary, aes(League, mean_rules, fill=League)) + 
  geom_col() +
  geom_errorbar(aes(ymin=mean_rules-sd_rules, ymax=mean_rules+sd_rules), width=0.2)

# group (supp)    score  (len)    spirit_cat (dose)
# m                 z                 rules
# w                 z                 rules
# x                 z                 rules

```

4. Who comments most: Men, Women, or Mixed?


3. Do development/second teams have better or worst spirit overall? 
I suspect for this one, they will be ranked lower because of rules and knowledge. That gives me a bit of a headache, because part of our job is to teach beginners the rules, and I'm not fussed if that happens at tournaments. To me, a well-spirited game with a beginner could go something like this:
*Pick Call*
Beginner: "What does that mean?"
Experienced Player: "It means your defender was stopped from defending you because they were close to you and you ran through people. Play stops and they get to catch up. Everybody goes back to where they were before the pick was called."
Beginner: "Oh, ok."
Doesn't seem like others feel that way though... 
```{r}
# for this use case, I only used teams 3 and 4 in my definition of "development" as many unis are big enough
# to easily field 2 "first" indoor teams.
dev_teams <- raw %>% 
  filter(grepl(4,Opponent) | grepl(3,Opponent))

experienced_teams <- raw %>%
  filter(grepl(1,Opponent))

compare_two_spirit_groups <- function(group1, group2, domain) {
  g1_domain <- group1[[domain]]
  g2_domain <- group2[[domain]]
  print(mean(g1_domain))
  print(mean(g2_domain))
  t.test(g1_domain, g2_domain)
}
```

Now actually compare them all
```{r}
compare_two_spirit_groups(dev_teams, experienced_teams, "Rules")
compare_two_spirit_groups(dev_teams, experienced_teams, "Fouls")
compare_two_spirit_groups(dev_teams, experienced_teams, "Fairness")
compare_two_spirit_groups(dev_teams, experienced_teams, "Attitude")
compare_two_spirit_groups(dev_teams, experienced_teams, "Comms")
```
So rules-wise, beginners teams have a harder time. Maybe we should go a little easier on them, or at least recognize it's our responsibility to teach them the rules. 
Everybody is equally fouly, lol.
Everybody is equally spirited when it comes to fairness.
Everybody seems to have a pretty high attitude.
And also everybody seems to be able to communicate pretty well.

Now let's see for men and women's
```{r}
compare_two_spirit_groups(men, women, "Rules")
compare_two_spirit_groups(men, women, "Fouls")
compare_two_spirit_groups(men, women, "Fairness")
compare_two_spirit_groups(men, women, "Attitude")
compare_two_spirit_groups(men, women, "Comms")
```
Everybody is equally bad at rules. 
Equally fouly
Equally fair
Equally good attitudes
But women rate each other lower on communication! zoinks

What about both of these but vs. mixed?
```{r}
compare_two_spirit_groups(men, mixed, "Rules")
compare_two_spirit_groups(men, mixed, "Fouls")
compare_two_spirit_groups(men, mixed, "Fairness")
compare_two_spirit_groups(men, mixed, "Attitude")
compare_two_spirit_groups(men, mixed, "Comms")


compare_two_spirit_groups(women, mixed, "Rules")
compare_two_spirit_groups(women, mixed, "Fouls")
compare_two_spirit_groups(women, mixed, "Fairness")
compare_two_spirit_groups(women, mixed, "Attitude")
compare_two_spirit_groups(women, mixed, "Comms")
```
The takeaway here is mixed gets round-the-board better spirit scores for everything. Playing mixed gender seems to make people happy.

4. On that note, who gets called out for contact most: Open, Women's, or Mixed? 

2. What part of the UK has the best/worst spirit (hehehe)? 
The weird thing about this is... it's self-reported. 

